The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2024-02-14 r_r/manuscript_rihe_r_r.tex
changed the introduction per reviewer 2

@@

\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs

@@

  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}

@@

  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}

@@

\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}

@@

\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\emph{During the 2017 calendar year, we tracked off-campus recruiting visits made by 14 selective private universities, 12 selective private liberal arts colleges, and 15 public research universities.\footnote{These institutions are: Northwestern, Notre Dame, Emory, Tufts, Boston College, Tulane, Case Western Reserve, Villanova, Southern Methodist University, Baylor, University of Denver, Texas Christian University, Stevens Institute of Technology, Marquette, Williams, Swarthmore, Middlebury, Smith, Harvey Mudd, Colorado College, Macalester, Scripps, Oberlin, Occidental, Sewanee, Connecticut College, UC-Berkeley, UC-San Diego, University of Georgia, University of Pittsburgh, Rutgers, University of Massachusetts-Amherst, UC-Riverside, SUNY Stony Brook, University of Colorado-Boulder, University of South Carolina, University of Kansas, University of Nebraska-Lincoln, University of Alabama, University of Cincinnati, and University of Arkansas.} We are still trying to figure out what happened.}

\bigskip

Autumn is college recruiting visit season at Harpeth Hall, a nonsectarian private school for girls located in the affluent Green Hills neighborhood of Nashville, Tennessee. Founded in 1951, Harpeth is an ``A+'' school ranked 337 in the nation according to the 2020 \emph{Best Private High Schools in America} by Niche. Enrollment patterns at Nashville-area private schools reflect broader regional patterns, whereby ``efforts by the courts to desegregate public schools in the South resulted in White flight to private schools'' (Murnane \& Reardon, 2018, p. 14). In Fall 2017, about 85\% of Harpeth students were white and less than 5\% were Black or Latinx (Author calculations).

From September 6th through October 17th, Harpeth received recruiting visits from 11 private universities in our sample (Northwestern, Boston College, Tufts, Villanova, Case Western Reserve, University of Denver, Emory, Tulane, Baylor, Texas Christian University, Southern Methodist University), and three private liberal arts colleges (Harvey Mudd, Scripps, and Sewanee). Previous scholarship observes strong connections between private schools and selective private universities (Khan, 2011; Stevens, 2007). However, Harpeth also received visits from six out-of-state public research universities, the University of Cincinnati, University of South Carolina, the University of Alabama, University of Arkansas, University of Georgia, and CU Boulder.

@@

Father Ryan High School is a Catholic school founded in 1925. Father Ryan is predominantly white (86\%, 7\% Black or Latinx) and received an ``A'' grade from Niche, meaning it is not ranked in the top 1,000 U.S. private schools. In Fall 2017, Father Ryan received visits from eight of the 14 private universities in our sample (including three of the four Catholic universities) and the same six public universities that visited Harpeth Hall, including two visits by the University of Alabama. Meanwhile, Christ Presbyterian Academy -- also not ranked in the top 1,000 private schools -- received visits from six private universities in our sample (including the three Christian universities Baylor, Texas Christian University, and Southern Methodist University). It also received visits from the University of Cincinnati, the University of South Carolina, the University of Alabama, University of Arkansas, and CU Boulder. Christ Presbyterian Academy was founded in 1985 and 94\% of its students are white.

Theories of organizational behavior suggest that university recruiting behaviors -- those that require significant expenditure -- are indicators of organizational enrollment priorities. Large, public-facing organizations are beset by multiple, competing pressures from the external environment, but organizations have difficulty pursuing multiple goals (Pfeffer \& Salancik, 1978; Thompson, 1967). Thus, organizations respond by symbolically adopting some goals and substantively adopting others (Meyer \& Rowan, 1977). Under symbolic adoption, organizations adopt policies and structures that send signals to external constituents but do not substantively affect resource allocation inside the ``technical level'' of the organization.\footnote{Contingency theory states that organizations consist of ``technical levels'' responsible for processing inputs and producing outputs and ``managerial levels'' responsible for directing the technical level and tending to demands from the external environment (Thompson, 1967). The technical level achieves effectiveness by focusing on a small, stable set of goals.} Under substantive adoption, organizations direct resources in the technical level towards achieving the goal. Off-campus recruiting visits to local high schools and college fairs was the top-ranked recruiting expenditure item for both public and private universities, according to Ruffalo Noel-Levitz (2020a). Therefore, analyses of which schools a university visits -- and which schools they ignore -- can yield insights about university enrollment goals.

Author (XXXXa) analyzed recruiting visits to public high schools made by 15 public research universities in 2017. The authors found that 12 of the 15 universities made more visits to out-of-state schools than in-state schools. These out-of-state visits focused on affluent, predominantly white public schools. Set against the scholarship about declining state appropriations and growing nonresident enrollment (Author, XXXXb, XXXXc), these results suggest that many public research universities prioritize enrollment from wealthy out-of-state students who can afford to pay nonresident tuition prices.

Within scholarship on social mobility and social stratification, a multidisciplinary literature analyzes the relationship between private school and college access (Bryk, Lee, \& Holland, 1993; Chetty, Deming, \& Friedman, 2023; Khan, 2011; Kingston \& Lewis, 1990). This research observes strong links -- and student flows -- between private high schools and selective private universities. Private schools enroll many students who can afford private university tuition (Stevens, 2007). To date, scholarship on nonresident enrollment by public universities has progressed in isolation from scholarship on links between private schools and colleges. We argue that these two stories have become connected because public research universities increasingly target out-of-state private schools, schools that have been the domain of selective private universities. Simply put, selective private universities and public research universities are increasingly competing for students from the same set of affluent households.

@@

This manuscript yields practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the Covid-19 Pandemic, off-campus recruiting visits have bounced back because they remain a fundamental means of maintaining relationships with high school guidance counselors (EAB \& NACAC, 2023; Ruffalo Noel-Levitz, 2022a, 2022b). Prior research finds that students are sensitive to recruiting interventions, including off-campus recruiting visits (Holland, 2019; Ruffalo Noel-Levitz, 2018). Therefore, which high schools a university chooses to visit affects the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race-conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The U.S. Department of Education (2023, p. 9) recommends that universities expand ``outreach to high schools with substantial populations of low-income students and students of color.'' Most public and private universities in our sample visited a disproportionate number of private schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

\hypertarget{nonresident-enrollment-at-public-universities}{%
\subsection{Nonresident Enrollment at Public Universities}\label{nonresident-enrollment-at-public-universities}}

@@

\hypertarget{linkages-between-private-schools-and-private-universities}{%
\subsection{Linkages Between Private Schools and Private Universities}\label{linkages-between-private-schools-and-private-universities}}

@@

\hypertarget{synthesis}{%
\subsection{Synthesis}\label{synthesis}}

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-and-sample}{%
\subsection{Data and Sample}\label{data-and-sample}}

@@

\hypertarget{constructing-social-networks}{%
\subsection{Constructing Social Networks}\label{constructing-social-networks}}

@@

\hypertarget{analyses}{%
\subsection{Analyses}\label{analyses}}

@@

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{rq1-scale-of-visits}{%
\subsection{RQ1: Scale of visits}\label{rq1-scale-of-visits}}

@@

\hypertarget{rq2-overlapping-network-structure}{%
\subsection{RQ2: Overlapping network structure}\label{rq2-overlapping-network-structure}}

@@

\hypertarget{rq3-characteristics-of-visited-schools}{%
\subsection{RQ3: Characteristics of visited schools}\label{rq3-characteristics-of-visited-schools}}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

@@

\hypertarget{implications-for-scholarship}{%
\subsection{Implications for Scholarship}\label{implications-for-scholarship}}

@@

\hypertarget{implications-for-practice}{%
\subsection{Implications for Practice}\label{implications-for-practice}}

@@

\hypertarget{conflict-of-interest-statement}{%
\section{Conflict of Interest Statement}\label{conflict-of-interest-statement}}

@@

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-RN4758}{}}%
Author. (XXXXa).

\leavevmode\vadjust pre{\hypertarget{ref-list_biz}{}}%
Author. (XXXXe).

\leavevmode\vadjust pre{\hypertarget{ref-list_empirics}{}}%
Author. (XXXXf).

\leavevmode\vadjust pre{\hypertarget{ref-RN3753}{}}%
Author. (XXXXb).

\leavevmode\vadjust pre{\hypertarget{ref-RN3685}{}}%
Author. (XXXXc).

\leavevmode\vadjust pre{\hypertarget{ref-RN4290}{}}%
Author. (XXXXd).

\leavevmode\vadjust pre{\hypertarget{ref-RN4953}{}}%
Borgatti, S. P. (1998). Ego networks. Retrieved from \url{http://www.analytictech.com/networks/egonet.htm}

\leavevmode\vadjust pre{\hypertarget{ref-RN2457}{}}%
Borgatti, S. P. (2008). 2-mode concepts in social network analysis. In \emph{Encyclopedia of complexity and system science}.

\leavevmode\vadjust pre{\hypertarget{ref-RN786}{}}%
Bryk, A. S., Lee, V. E., \& Holland, P. B. (1993). \emph{Catholic schools and the common good}. Cambridge: Harvard University Press.

\leavevmode\vadjust pre{\hypertarget{ref-RN975}{}}%
Burt, R. S. (1992). \emph{Structural holes: The social structure of competition}. Cambridge, MA: Harvard University Press.

\leavevmode\vadjust pre{\hypertarget{ref-RN2772}{}}%
Cheslock, J. J., \& Kroc, R. (2012). Managing college enrollments. In R. Howard, B. Knight, \& G. McLaughlin (Eds.), \emph{The handbook for institutional researchers} (pp. 221--236). San Francisco, CA: Jossey-Bass.

\leavevmode\vadjust pre{\hypertarget{ref-RN4935}{}}%
Chetty, R., Deming, D. J., \& Friedman, J. N. (2023). Diversifying society's leaders? The causal effects of admission to highly selective private colleges. \emph{National Bureau of Economic Research Working Paper Series}, \emph{No. 31492}. \url{https://doi.org/10.3386/w31492}

\leavevmode\vadjust pre{\hypertarget{ref-RN4599}{}}%
Clark, D., \& Del Bono, E. (2016). The long-run effects of attending an elite school: Evidence from the united kingdom. \emph{American Economic Journal-Applied Economics}, \emph{8}(1), 150--176. \url{https://doi.org/10.1257/app.20130505}

\leavevmode\vadjust pre{\hypertarget{ref-RN4677}{}}%
Clauset, A., Newman, M. E. J., \& Moore, C. (2004). Finding community structure in very large networks. \emph{Physical Review E}, \emph{70}(6). \url{https://doi.org/10.1103/PhysRevE.70.066111}

\leavevmode\vadjust pre{\hypertarget{ref-RN4937}{}}%
Coleman, J. S., \& Hoffer, T. (1987). \emph{Public and private high schools: The impact of communities} (pp. xxviii, 254 p.). New York: Basic Books.

\leavevmode\vadjust pre{\hypertarget{ref-RN4647}{}}%
Cookson, P. W., \& Persell, C. H. (1985). \emph{Preparing for power : America's elite boarding schools} (pp. x, 260 p.). New York: Basic Books.

\leavevmode\vadjust pre{\hypertarget{ref-RN1791}{}}%
Curs, B. R., \& Singell, L. D. (2002). An analysis of the application and enrollment processes for in-state and out-of-state students at a large public university. \emph{Economics of Education Review}, \emph{21}(2), 111--124. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://000173740100002}{\textless Go to ISI\textgreater://000173740100002}

\leavevmode\vadjust pre{\hypertarget{ref-RN2287}{}}%
Curs, B. R., \& Singell, L. D. (2010). Aim high or go low? Pricing strategies and enrollment effects when the net price elasticity varis with need and ability. \emph{Journal of Higher Education}, \emph{81}(4), 515--543.

\leavevmode\vadjust pre{\hypertarget{ref-RN4663}{}}%
Davis, G. F., Yoo, M., \& Baker, W. E. (2003). The small world of the american corporate elite, 1982-2001. \emph{Strategic Organization}, \emph{1}(3), 301--326. \url{https://doi.org/10.1177/14761270030013002}

\leavevmode\vadjust pre{\hypertarget{ref-RN2271}{}}%
Delaney, J. A., \& Doyle, W. R. (2011). State spending on higher education: Testing the balance wheel over time. \emph{Journal of Education Finance}, \emph{36}(4), 343--368. Journal Article.

\leavevmode\vadjust pre{\hypertarget{ref-RN3524}{}}%
DesJardins, S. L., \& Bell, A. (2006). Using economic concepts to inform enrollment management. \emph{New Directions for Institutional Research}, \emph{2006}(132), 59--74. \url{https://doi.org/10.1002/ir.196}

\leavevmode\vadjust pre{\hypertarget{ref-eab_nacac_counselor_2023}{}}%
EAB, \& NACAC. (2023). \emph{The counselor perspective: Insights from EAB's survey of high school counselors}. EAB. Retrieved from \url{https://www.nacacnet.org/wp-content/uploads/The-Counselor-Perspective-EAB-Survey.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-RN4936}{}}%
Eide, E. R., Goldhaber, D. D., \& Showalter, M. H. (2004). Does catholic high school attendance lead to attendance at a more selective college? \emph{Social Science Quarterly}, \emph{85}(5), 1335--1352. \url{https://doi.org/10.1111/j.0038-4941.2004.00279.x}

\leavevmode\vadjust pre{\hypertarget{ref-RN4116}{}}%
Eisenhardt, K. M. (1989). Building theories from case study research. \emph{Academy of Management Review}, \emph{14}(4), 532--550. \url{https://doi.org/10.2307/258557}

\leavevmode\vadjust pre{\hypertarget{ref-RN4642}{}}%
Falsey, B., \& Heyns, B. (1984). The college channel: Private and public schools reconsidered. \emph{Sociology of Education}, \emph{57}(2), 111--122. \url{https://doi.org/10.2307/2112633}

\leavevmode\vadjust pre{\hypertarget{ref-RN4724}{}}%
Graham, L. (1999). \emph{Our kind of people: Inside america's black upper class} (1st ed., pp. xii, 418 pages, 16 unnumbered pages of plates). New York: HarperCollins.

\leavevmode\vadjust pre{\hypertarget{ref-RN4625}{}}%
Hill, L. D. (2008). School strategies and the "college-linking" process: Reconsidering the effects of high schools on college enrollment. \emph{Sociology of Education}, \emph{81}(1), 53--76. Journal Article. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://WOS:000256634400003}{\textless Go to ISI\textgreater://WOS:000256634400003}

\leavevmode\vadjust pre{\hypertarget{ref-RN4324}{}}%
Holland, M. M. (2019). \emph{Divergent paths to college: Race, class, and inequality in high schools}. Rutgers University Press. \url{https://doi.org/10.36019/9780813590288}

\leavevmode\vadjust pre{\hypertarget{ref-em_burnout_hoover}{}}%
Hoover, E. (2023). A profession on the edge: Why enrollment leaders are wearing down, burning out, and leaving jobs they once loved. \emph{The Chronicle of Higher Education}, (April 5). Retrieved from \url{https://www.chronicle.com/article/a-profession-on-the-edge}

\leavevmode\vadjust pre{\hypertarget{ref-RN2133}{}}%
Hoxby, C. M. (1997). \emph{How the changing market structure of u.s. Higher education explains college tuition} (No. Working Paper 6323). National Bureau of Economic Research. Retrieved from \url{http://www.nber.org/papers/W6323.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-RN4495}{}}%
Jack, A. A. (2019). \emph{The privileged poor: How elite colleges are failing disadvantaged students} (p. 276). Cambridge, Massachusetts: Harvard University Press,. Retrieved from \href{Restricted\%20to\%20UCLA.\%20Limited\%20to\%20three\%20concurrent\%20users.\%20Try\%20again\%20later\%20if\%20unavailable.\%20http://openurl.cdlib.org?sid=UCLA:CAT\&genre=book\&__char_set=utf8\&isbn=9780674239647}{Restricted to UCLA. Limited to three concurrent users. Try again later if unavailable. http://openurl.cdlib.org?sid=UCLA:CAT\&genre=book\&\_\_char\_set=utf8\&isbn=9780674239647}

\leavevmode\vadjust pre{\hypertarget{ref-RN4596}{}}%
Jerrim, J., Parker, P. D., Chmielewski, A. K., \& Anders, J. (2016). Private schooling, educational transitions, and early labour market outcomes: Evidence from three anglophone countries. \emph{European Sociological Review}, \emph{32}(2), 280--294. \url{https://doi.org/10.1093/esr/jcv098}

>>>>

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}

@@

\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}

@@

   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}

@@

  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}

@@

\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}

@@

\usepackage{bookmark}

@@

\section{Introduction}\label{introduction}

Autumn is college recruiting season at Harpeth Hall, a nonsectarian private school for girls located in the affluent Green Hills neighborhood of Nashville, Tennessee. Founded in 1951, Harpeth is an ``A+'' school ranked 337 in the nation according to the 2020 \emph{Best Private High Schools in America} by Niche. Enrollment patterns at Nashville-area private schools reflect broader regional patterns, whereby ``efforts by the courts to desegregate public schools in the South resulted in White flight to private schools'' (Murnane \& Reardon, 2018, p. 14). In Fall 2017, about 85\% of Harpeth students were white and less than 5\% were Black or Latinx (Author calculations).

During the 2017 calendar year, we tracked off-campus recruiting visits made by 14 selective private universities, 12 selective private liberal arts colleges, and 15 public research universities.\footnote{These institutions are: Northwestern, Notre Dame, Emory, Tufts, Boston College, Tulane, Case Western Reserve, Villanova, Southern Methodist University, Baylor, University of Denver, Texas Christian University, Stevens Institute of Technology, Marquette, Williams, Swarthmore, Middlebury, Smith, Harvey Mudd, Colorado College, Macalester, Scripps, Oberlin, Occidental, Sewanee, Connecticut College, UC-Berkeley, UC-San Diego, University of Georgia, University of Pittsburgh, Rutgers, University of Massachusetts-Amherst, UC-Riverside, SUNY Stony Brook, University of Colorado-Boulder, University of South Carolina, University of Kansas, University of Nebraska-Lincoln, University of Alabama, University of Cincinnati, and University of Arkansas.} From September 6th through October 17th, Harpeth Hall received recruiting visits from 11 private universities in our sample (Northwestern, Boston College, Tufts, Villanova, Case Western Reserve, University of Denver, Emory, Tulane, Baylor, Texas Christian University, Southern Methodist University), and three private liberal arts colleges (Harvey Mudd, Scripps, and Sewanee). Previous scholarship observes strong connections between private schools and selective private universities (Khan, 2011; Stevens, 2007). However, Harpeth also received visits from six out-of-state public research universities, the University of Cincinnati, University of South Carolina, the University of Alabama, University of Arkansas, University of Georgia, and CU Boulder.

@@

Christ Presbyterian Academy, located in the Forest Hills suburb southwest of Nashville, received an ``A'' grade from Niche, meaning it is not ranked in the top 1,000 U.S. private schools. In 2017, it received visits from six private universities in our sample (including the three Christian universities, Baylor, Texas Christian University, and Southern Methodist University). It also received visits from the University of Cincinnati, the University of South Carolina, the University of Alabama, University of Arkansas, and CU Boulder. Christ Presbyterian Academy was founded in 1985 and 94\% of its students are white.

This paper analyzes the social networks revealed by recruiting visits from universities to private high schools. Theories of organizational behavior suggest that university recruiting behaviors -- those that require significant expenditure -- are indicators of organizational enrollment priorities. Public-facing organizations are beset by multiple, competing pressures from the external environment, but organizations have difficulty pursuing multiple goals (Pfeffer \& Salancik, 1978; Thompson, 1967). Thus, organizations respond by symbolically adopting some goals and substantively adopting others (Meyer \& Rowan, 1977). Under symbolic adoption, organizations adopt policies and structures that send signals to external constituents but do not substantively affect resource allocation inside the ``technical level'' of the organization.\footnote{Contingency theory states that organizations consist of ``technical levels'' responsible for processing inputs and producing outputs and ``managerial levels'' responsible for directing the technical level and tending to demands from the external environment (Thompson, 1967). The technical level achieves effectiveness by focusing on a small, stable set of goals.} Under substantive adoption, organizations direct resources in the technical level towards achieving the goal. Off-campus recruiting visits to local high schools and college fairs was the top-ranked recruiting expenditure item for both public and private universities, according to Ruffalo Noel-Levitz (2020a). Therefore, analyses of which schools a university visits -- and which schools they ignore -- can yield insights about university enrollment goals.

We situate this study within scholarship on nonresident enrollment at public universities. Author (XXXXa) analyzed recruiting visits to public high schools made by 15 public research universities in 2017. The authors found that 12 of the 15 universities made more visits to out-of-state schools than in-state schools. These out-of-state visits focused on affluent, predominantly white public schools. Set against the scholarship about declining state appropriations and growing nonresident enrollment (Author, XXXXb, XXXXc), these results suggest that many public research universities prioritize enrollment from wealthy out-of-state students who can afford to pay nonresident tuition prices.

A separate, multidisciplinary literature has analyzed linkages betwen private school and college access (Bryk, Lee, \& Holland, 1993; Chetty, Deming, \& Friedman, 2023; Khan, 2011; Kingston \& Lewis, 1990). This research observes strong links -- and student flows -- between private high schools and selective private universities. Private schools enroll many students who can afford private university tuition (Stevens, 2007). To date, scholarship on nonresident enrollment by public universities has progressed in isolation from scholarship on links between private schools and colleges.

We argue that these two literatures have become connected because public research universities increasingly target out-of-state private schools, schools that have been the domain of selective private universities. Simply put, selective private universities and public research universities are increasingly competing for students from the same set of affluent households.

@@

The results yield practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the Covid-19 Pandemic, off-campus recruiting visits have bounced back because they remain a fundamental means of maintaining relationships with high school guidance counselors (EAB \& NACAC, 2023; Ruffalo Noel-Levitz, 2022a, 2022b). Prior research finds that students are sensitive to recruiting interventions, including off-campus recruiting visits (Holland, 2019; Ruffalo Noel-Levitz, 2018). Therefore, which high schools a university chooses to visit affects the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race-conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The U.S. Department of Education (2023, p. 9) recommends that universities expand ``outreach to high schools with substantial populations of low-income students and students of color.'' Most public and private universities in our sample visited a disproportionate number of private schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

\section{Literature Review}\label{literature-review}

\subsection{Nonresident Enrollment at Public Universities}\label{nonresident-enrollment-at-public-universities}

@@

\subsection{Linkages Between Private Schools and Private Universities}\label{linkages-between-private-schools-and-private-universities}

@@

\subsection{Synthesis}\label{synthesis}

@@

\section{Methods}\label{methods}

\subsection{Data and Sample}\label{data-and-sample}

@@

\subsection{Constructing Social Networks}\label{constructing-social-networks}

@@

\subsection{Analyses}\label{analyses}

@@

\subsection{Limitations}\label{limitations}

@@

\section{Results}\label{results}

\subsection{RQ1: Scale of visits}\label{rq1-scale-of-visits}

@@

\subsection{RQ2: Overlapping network structure}\label{rq2-overlapping-network-structure}

@@

\subsection{RQ3: Characteristics of visited schools}\label{rq3-characteristics-of-visited-schools}

@@

\section{Discussion}\label{discussion}

\subsection{Summary}\label{summary}

@@

\subsection{Implications for Scholarship}\label{implications-for-scholarship}

@@

\subsection{Implications for Practice}\label{implications-for-practice}

@@

\section{Conflict of Interest Statement}\label{conflict-of-interest-statement}

@@

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-RN4758}
Author. (XXXXa).

\bibitem[\citeproctext]{ref-list_biz}
Author. (XXXXe).

\bibitem[\citeproctext]{ref-list_empirics}
Author. (XXXXf).

\bibitem[\citeproctext]{ref-RN3753}
Author. (XXXXb).

\bibitem[\citeproctext]{ref-RN3685}
Author. (XXXXc).

\bibitem[\citeproctext]{ref-RN4290}
Author. (XXXXd).

\bibitem[\citeproctext]{ref-RN4953}
Borgatti, S. P. (1998). Ego networks. Retrieved from \url{http://www.analytictech.com/networks/egonet.htm}

\bibitem[\citeproctext]{ref-RN2457}
Borgatti, S. P. (2008). 2-mode concepts in social network analysis. In \emph{Encyclopedia of complexity and system science}.

\bibitem[\citeproctext]{ref-RN786}
Bryk, A. S., Lee, V. E., \& Holland, P. B. (1993). \emph{Catholic schools and the common good}. Cambridge: Harvard University Press.

\bibitem[\citeproctext]{ref-RN975}
Burt, R. S. (1992). \emph{Structural holes: The social structure of competition}. Cambridge, MA: Harvard University Press.

\bibitem[\citeproctext]{ref-RN2772}
Cheslock, J. J., \& Kroc, R. (2012). Managing college enrollments. In R. Howard, B. Knight, \& G. McLaughlin (Eds.), \emph{The handbook for institutional researchers} (pp. 221--236). San Francisco, CA: Jossey-Bass.

\bibitem[\citeproctext]{ref-RN4935}
Chetty, R., Deming, D. J., \& Friedman, J. N. (2023). Diversifying society's leaders? The causal effects of admission to highly selective private colleges. \emph{National Bureau of Economic Research Working Paper Series}, \emph{No. 31492}. \url{https://doi.org/10.3386/w31492}

\bibitem[\citeproctext]{ref-RN4599}
Clark, D., \& Del Bono, E. (2016). The long-run effects of attending an elite school: Evidence from the united kingdom. \emph{American Economic Journal-Applied Economics}, \emph{8}(1), 150--176. \url{https://doi.org/10.1257/app.20130505}

\bibitem[\citeproctext]{ref-RN4677}
Clauset, A., Newman, M. E. J., \& Moore, C. (2004). Finding community structure in very large networks. \emph{Physical Review E}, \emph{70}(6). \url{https://doi.org/10.1103/PhysRevE.70.066111}

\bibitem[\citeproctext]{ref-RN4937}
Coleman, J. S., \& Hoffer, T. (1987). \emph{Public and private high schools: The impact of communities} (pp. xxviii, 254 p.). New York: Basic Books.

\bibitem[\citeproctext]{ref-RN4647}
Cookson, P. W., \& Persell, C. H. (1985). \emph{Preparing for power : America's elite boarding schools} (pp. x, 260 p.). New York: Basic Books.

\bibitem[\citeproctext]{ref-RN1791}
Curs, B. R., \& Singell, L. D. (2002). An analysis of the application and enrollment processes for in-state and out-of-state students at a large public university. \emph{Economics of Education Review}, \emph{21}(2), 111--124. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://000173740100002}{\textless Go to ISI\textgreater://000173740100002}

\bibitem[\citeproctext]{ref-RN2287}
Curs, B. R., \& Singell, L. D. (2010). Aim high or go low? Pricing strategies and enrollment effects when the net price elasticity varis with need and ability. \emph{Journal of Higher Education}, \emph{81}(4), 515--543.

\bibitem[\citeproctext]{ref-RN4663}
Davis, G. F., Yoo, M., \& Baker, W. E. (2003). The small world of the american corporate elite, 1982-2001. \emph{Strategic Organization}, \emph{1}(3), 301--326. \url{https://doi.org/10.1177/14761270030013002}

\bibitem[\citeproctext]{ref-RN2271}
Delaney, J. A., \& Doyle, W. R. (2011). State spending on higher education: Testing the balance wheel over time. \emph{Journal of Education Finance}, \emph{36}(4), 343--368. Journal Article.

\bibitem[\citeproctext]{ref-RN3524}
DesJardins, S. L., \& Bell, A. (2006). Using economic concepts to inform enrollment management. \emph{New Directions for Institutional Research}, \emph{2006}(132), 59--74. \url{https://doi.org/10.1002/ir.196}

\bibitem[\citeproctext]{ref-eab_nacac_counselor_2023}
EAB, \& NACAC. (2023). \emph{The counselor perspective: Insights from EAB's survey of high school counselors}. EAB. Retrieved from \url{https://www.nacacnet.org/wp-content/uploads/The-Counselor-Perspective-EAB-Survey.pdf}

\bibitem[\citeproctext]{ref-RN4936}
Eide, E. R., Goldhaber, D. D., \& Showalter, M. H. (2004). Does catholic high school attendance lead to attendance at a more selective college? \emph{Social Science Quarterly}, \emph{85}(5), 1335--1352. \url{https://doi.org/10.1111/j.0038-4941.2004.00279.x}

\bibitem[\citeproctext]{ref-RN4116}
Eisenhardt, K. M. (1989). Building theories from case study research. \emph{Academy of Management Review}, \emph{14}(4), 532--550. \url{https://doi.org/10.2307/258557}

\bibitem[\citeproctext]{ref-RN4642}
Falsey, B., \& Heyns, B. (1984). The college channel: Private and public schools reconsidered. \emph{Sociology of Education}, \emph{57}(2), 111--122. \url{https://doi.org/10.2307/2112633}

\bibitem[\citeproctext]{ref-RN4724}
Graham, L. (1999). \emph{Our kind of people: Inside america's black upper class} (1st ed., pp. xii, 418 pages, 16 unnumbered pages of plates). New York: HarperCollins.

\bibitem[\citeproctext]{ref-RN4625}
Hill, L. D. (2008). School strategies and the "college-linking" process: Reconsidering the effects of high schools on college enrollment. \emph{Sociology of Education}, \emph{81}(1), 53--76. Journal Article. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://WOS:000256634400003}{\textless Go to ISI\textgreater://WOS:000256634400003}

\bibitem[\citeproctext]{ref-RN4324}
Holland, M. M. (2019). \emph{Divergent paths to college: Race, class, and inequality in high schools}. Rutgers University Press. \url{https://doi.org/10.36019/9780813590288}

\bibitem[\citeproctext]{ref-em_burnout_hoover}
Hoover, E. (2023). A profession on the edge: Why enrollment leaders are wearing down, burning out, and leaving jobs they once loved. \emph{The Chronicle of Higher Education}, (April 5). Retrieved from \url{https://www.chronicle.com/article/a-profession-on-the-edge}

\bibitem[\citeproctext]{ref-RN2133}
Hoxby, C. M. (1997). \emph{How the changing market structure of u.s. Higher education explains college tuition} (No. Working Paper 6323). National Bureau of Economic Research. Retrieved from \url{http://www.nber.org/papers/W6323.pdf}

\bibitem[\citeproctext]{ref-RN4495}
Jack, A. A. (2019). \emph{The privileged poor: How elite colleges are failing disadvantaged students} (p. 276). Cambridge, Massachusetts: Harvard University Press,. Retrieved from \href{Restricted\%20to\%20UCLA.\%20Limited\%20to\%20three\%20concurrent\%20users.\%20Try\%20again\%20later\%20if\%20unavailable.\%20http://openurl.cdlib.org?sid=UCLA:CAT&genre=book&__char_set=utf8&isbn=9780674239647}{Restricted to UCLA. Limited to three concurrent users. Try again later if unavailable. http://openurl.cdlib.org?sid=UCLA:CAT\&genre=book\&\_\_char\_set=utf8\&isbn=9780674239647}

\bibitem[\citeproctext]{ref-RN4596}
Jerrim, J., Parker, P. D., Chmielewski, A. K., \& Anders, J. (2016). Private schooling, educational transitions, and early labour market outcomes: Evidence from three anglophone countries. \emph{European Sociological Review}, \emph{32}(2), 280--294. \url{https://doi.org/10.1093/esr/jcv098}

\bibitem[\citeproctext]{ref-RN4321}
Khan, S. R. (2010). \emph{Getting in: How elite schools play the college game} (pp. 97--113). Rowman \& Littlefield.

\bibitem[\citeproctext]{ref-RN4407}
Khan, S. R. (2011). \emph{Privilege: The making of an adolescent elite at st. Paul's school} (pp. 232 pages). Princeton, N.J.: Princeton University Press.

\bibitem[\citeproctext]{ref-RN3522}
Killgore, L. (2009). Merit and competition in selective college admissions. \emph{Review of Higher Education}, \emph{32}(4), 469--488. Retrieved from \href{\%3CGo\%20to\%20ISI\%3E://WOS:000266737500002}{\textless Go to ISI\textgreater://WOS:000266737500002}

\bibitem[\citeproctext]{ref-RN1565}
Kingston, P. W., \& Lewis, L. S. (1990). \emph{The high-status track: Studies of elite schools and stratification} (pp. xxxiv, 261 p.). Albany: State University of New York Press.
